Errorless irrationality: removing error-driven components from the inverse base-rate effect paradigm

### 2023-01-03 docs/manuscript.tex
expand related work and add results

%\setlength\titlebox{4.5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 4.5cm (the original size).
%%If you do, we reserve the right to require you to change it back in
%%the camera-ready version, which could interfere with the timely
%%appearance of your paper in the Proceedings.

\title{Errorless irrationality: removing prediction errors from the inverse base-rate effect}

@@

We will build on work carried out by \citeA{johansen2007paradoxical}.
\citeA{johansen2007paradoxical} failed to observe the inverse base-rate effect in an observational learning condition, where category labels were presented with features at the same time to participants.
They largely interpreted their results in opposition to EXIT \cite{kruschke2001toward}, but their experimental procedure does not correspond to the standard inverse base-rate effect procedure.
Their design involves disjoint-cues (where categories shared no features in common), while the canonical design depends on a shared features during training that facilitates attentional relocation, which in turn moves responding towards the irrational rare preference.
The only instance when they observed a rare bias was when common features presented in compound during training were paired with a rare feature presented by itself during training.
A simple explanation for that result is that the rare feature developed stronger connections with the category label.
Compound features share the connection with the category label, so any update to these connections dissipate between them.
There is no need to relocate attention to reduce errors, therefore attention will not push responding towards the rare label.
Their design is therefore not sufficient to answer our question, whether prediction error is a necessary condition.

@@

\begin{table}[!ht]
  \begin{center}
    \caption{Abstract design of Experiment 1 including both test and training phases. \\}
    \label{tab:abstract-exp1}
    \begin{tabular}{llr} % text alignments
      \textbf{Training (Relative Frequencies)} & \textbf{Test}& \\
      \hline
      & \\
      $AB \to common_{1}$ (x 3) &  A, B, C,         &  \\
      $AC \to rare_{1}$   (x 1) &  AB, AC, BC      & x 20 \\
      \hline
    \end{tabular}
  \end{center}
\end{table}

@@

@@

\begin{table}[!ht]
  \begin{center}
    \caption{Abstract design of Experiment 2 including both test and training phases. X and Y are in place of the category labels. During the test phase, participants needed to select either X or Y to complete the features shown below.\\}
    \label{tab:abstract-exp2}
    \begin{tabular}{llr} % text alignments
      \textbf{Training (Relative Frequencies)} & \textbf{Test}& \\
      \hline
      & \\
      ABX x 3 &  A, B, C,         &  \\
      ACY x 1 &  AB, AC, BC      & x 20 \\
      \hline
    \end{tabular}
  \end{center}
\end{table}

@@

>>>>

%\setlength\titlebox{4.5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 4.5cm (the original size).
%%If you do, we reserve the right to require you to change it back in
%%the camera-ready version, which could interfere with the timely
%%appearance of your paper in the Proceedings.
\def\arraystretch{1.15}%  1 is the default, change whatever you need

\title{Errorless irrationality: removing error-driven components from the inverse base-rate effect paradigm}

@@

To our knowledge, there is only one attempt to implement the IBRE procedure without explicit feedback.
\citeA{johansen2007paradoxical} tried to observe the inverse base-rate effect both in a predictive-learning condition and in an observational learning condition, where category labels were presented with features at the same time to participants.
Their design involves disjoint-cues (where categories shared no features in common), while the canonical design depends on a shared feature during training that facilitates attentional relocation.
This attentional tuning in turn pushes responding towards the rare label.
Their design was optimised to investigate the hypothesised assymetric cognitive representation of the two categories - one of the assumptions.
As a result, the only instance when they observed a rare bias was when common features presented in compound during training were paired with a rare feature presented by itself during training.
This provided evidence for the assymetric cognitive represetation hypothesised to develop during training.
In order to investigate the role of prediction error in response to feedback, we need to tweak their observational learning condition to conform to a more canonical implementation of the procedure.
A simple neural-network explanation for the rare bias \citeA{johansen2007paradoxical} observed is that the rare feature developed stronger connections with the category label.
Compound features share the connection with the category label, so any update to these connections dissipate between them.
There is no need to relocate attention to reduce errors, therefore attention will not bias responding towards the rare label.

The only two studies directly looking at error-driven processes in the IBRE are \citeA{inkster2022neural} and \citeA{wills2014attention}.
\citeA{inkster2022neural} carried out a direct investigation into brain regions underlying error-driven learning in the IBRE.
They observed

@@

\begin{table}[!ht]
  \begin{center}
    \caption{Abstract design of Experiment 1 including both test and training phases. \\}
    \label{tab:abstract-exp1}
    \begin{tabular}{llr} % text alignments
      \textbf{Training (Relative Frequencies)} & \textbf{Test}& \\
      \hline
      % & \\
      $AB \to common_{1}$ (x 3) &  A, B, C,         &  \\
      $AC \to rare_{1}$   (x 1) &  AB, AC, BC      & x 20 \\
      \hline
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table}[H]
  \begin{center}
    \caption{Stuff. \\}
    \label{tab:results-exp1}
    \vskip 0.12in
    \begin{tabular}{rcc}
      \hline
      & $P(common)$ & $P(rare)$ \\
      \hline
      A & 0.69 & 0.31\\
      AB & 0.93 & 0.07\\
      AC & 0.10 & 0.90\\
      B & 0.93 & 0.07\\
      \textbf{BC} & \textbf{0.33} & \textbf{0.66}\\
      C & 0.06 & 0.94\\
    \end{tabular}
  \end{center}
\end{table}

$M = 0.66$, 95\% HDI $[0.62, 0.71]$, $\mathrm{BF}_{10} = 6.63 \times 10^{7}$

@@

\begin{table}[!ht]
  \begin{center}
    \caption{Abstract design of Experiment 2 including both test and training phases. X and Y are in place of the category labels. During the test phase, participants needed to select either X or Y to complete the features shown below.\\}
    \label{tab:abstract-exp2}
    \begin{tabular}{llr} % text alignments
      \textbf{Training (Relative Frequencies)} & \textbf{Test}& \\
      \hline
      % & \\
      ABX x 3 &  A, B, C,         &  \\
      ACY x 1 &  AB, AC, BC      & x 20 \\
      \hline
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table}[H]
  \begin{center}
    \caption{Caption.\\}
    \label{tab:results-exp2}
    \vskip 0.12in
    \begin{tabular}{rcc}
      \hline
       & $P(common)$ & $P(rare)$ \\
      \hline
      A & 0.76 & 0.24\\
      AB & 0.93 & 0.07\\
      AC & 0.11 & 0.89\\
      B & 0.92 & 0.08\\
      \textbf{BC} & \textbf{0.34} & \textbf{0.65}\\
      C & 0.09 & 0.91\\
    \end{tabular}
  \end{center}
\end{table}

$M = 0.65$, 95\% HDI $[0.55, 0.76]$, $\mathrm{BF}_{10} = 8.13$
